\capitulo{2}{Project objectives}
The general objective of this work is to design and develop an automatic music style recognition system using artificial intelligence. 
To achieve this general objective, the following specific objectives are proposed:

\subsection{Data collection}
The process of data collection involves gathering and organizing various types of information related to the data that we want to study. In this project, this include audio tracks and metadata obtained from FMA (Free Music Archive) dataset.
\subsection{Data preprocessing}
Preprocessing data in a musical dataset involves a series of steps aimed at cleaning, transforming, and preparing the data for further analysis or modeling tasks.
\subsection{Data processing}
Data processing in this project means extract relevant features to get relevant information to perform the classification. Features like MFCC, Spectogram  or Chromagrams are suitable for this task.
\subsection{Selection and implementation of models}
The selection and implementation of models involve choosing appropriate machine learning algorithms and techniques for music style recognition.
\subsection{Model evaluation}
Model evaluation involves measuring the performance and effectiveness of the implemented machine learning models with metrics like Accuracy, Precision or F1 Score.
\subsection{User interface}
The development of a user interface allows the user to interact with the music recognition model in an easy way.
\subsection{Deployment}
Deploy the web application on a web server or a cloud platform to ensure that is accessible to users.
\subsection{Testing and quality assurance}
Testing the application and web interface to identify and fix any bugs or issues. This process is crucial in the entire life-cycle of the application and development process.